\thispagestyle{empty}
\begin{raggedleft}
\vspace*{23mm}
\hfill {\huge {\bf {Acknowledgements}}} \\
\vspace{6mm}
\hfill \rule{4in}{.015in} \\
\vspace{19mm}
\end{raggedleft}


\noindent{There are so many people that I want to thank -- you all, I hope, know who you are. But the last three years have been eventful and the writing of this thesis has been one of the lesser challenges.  
%Anyone reading this will probably understand what it means when I say that the writing of this thesis has been among the lesser challenges that I have faced during my PhD.  
And so, there are just a few, very heartfelt and uncharacteristically emotional, thanks that I wish to make.}
%\bigskip
%\noindent 
First and foremost, to Mike;  you have been endlessly patient and have always been generous with your time and advice.  You gave me the  space to find my own path through my PhD, particularly in difficult times, and I am so grateful for this.  
%You have gone above and beyond, especially in difficult times, and I truly thank you.
To Jeremy; %your support over the last three years has been invaluable. ]
you advised me thoughtfully and carefully. Thanks for always having  encouraging words at the right moments.
%  your encouragement always seemed to come at the right times.  
To Patrick; %simply put, this thesis would not exist without you!  
thank you for everything -- %for advice on science and on life, for gossipy drinks and endless cups of tea, for giggling from the start to the end, for your patient understanding and for tea gin. 
from science advice to gossip to tea gin, this thesis probably wouldn't exist without you!
%\bigskip 
%\noindent 
To 
Benedict,
Dom,
Fiona,
Georgia,
George,
Jack,
and
Mike;  my luck in counting you as my friends is beyond words. %I know that every one of you would be there if I needed you, and 
Thank you for the
% proved this on numerous occasions over the last three years.  
%My life is so much the better for having the 
conversation, laughter and wisdom that you all bring to my life.  
%\clearpage
%\bigskip
%\noindent 
To Alex;  good job growing up! I could not be prouder to call you my brother.  Knowing that you are always there for me means so much. 
%\bigskip
%
%\noindent 
To Nick; you have been there for me in my lowest moments and my best.  
I cannot imagine having done this without you, and I have so many memories that I will cherish for the rest of my life.  
You got me through this 
%, as you get me through so much of my life.  
and I am so grateful.
%\bigskip
%\noindent 
To Mum and Dad;  you have been, are, and will always be my best friends in the world.  Support does not even begin to cover what you have given me, not just over the last three years but my whole life.  You are both caring, generous and wise, and I am so lucky to be able to call you Mum and Dad.  
\bigskip

\noindent{Thank you.}




%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% If you have more than one page in your acknowledgements, include the
% following on each page after the first:
%
% \thispagestyle{fancyplain}
% \lhead[\fancyplain{}{\vspace{4mm}\thepage \vspace{0mm}}]{}
% \rhead[]{\fancyplain{}{\vspace{4mm}\thepage \vspace{0mm}}}
% \cfoot{}
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

